### 2013-02-08 paper/ms.tex
including new tables with a summary

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJ Letters}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}
%\bibliographystyle{apj}
\usepackage{graphics}
\usepackage[dvips]{epsfig}

\newcommand{\kms}{\,km~s$^{-1}$}
\def\squig{\sim\!\!}
\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\Lsun}{\mbox{\,$L_{\odot}$}}
\newcommand{\like}{\mathscr{L}}
\newcommand{\bftheta}{\mathbf{\Theta}}
\newcommand{\degree}{\ensuremath{^\circ}}
\def\spose#1{\hbox to 0pt{#1\hss}}
\def\simlt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13C$}}}
\def\simgt{\mathrel{\spose{\lower 3pt\hbox{$\mathchar"218$}}
     \raise 2.0pt\hbox{$\mathchar"13E$}}}
\font\smcap=cmcsc10
\newcommand{\YH}[1]{\textcolor{blue}{\bf YH: #1}}
\newcommand{\NEW}[1]{\textcolor{blue}{  #1}}
\newcommand{\SOUT}[1]{\textcolor{red}{\sout{  #1}}}

\shorttitle{Local Group Kinematics}
\shortauthors{Forero-Romero et al.}

\begin{document}
\title{The kinematics of the Local Group in a cosmological context}
\author{
J. E.\ Forero-Romero\altaffilmark{1}, 
Y. Hoffman\altaffilmark{2}, 
S. Bustamante\altaffilmark{3}, 
S. Gottl\"ober\altaffilmark{4}, 
G. Yepes\altaffilmark{5}
}

\altaffiltext{1}{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia, \email{je.forero@uniandes.edu.co}}
\altaffiltext{2}{Racah Institute of Physics, The Hebrew University of Jerusalem, 91904 Jerusalem, Israel}
\altaffiltext{3}{Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle 67 No. 53-108, Medell\'{\i}n, Colombia}
\altaffiltext{4}{Leibniz-Institut f\"ur Astrophysik, Potsdam, An der Sternwarte 16, 14482 Potsdam, Germany}
\altaffiltext{5}{Grupo de Astrof\'{\i}sica, Departamento de F\'{\i}sica Te\'orica, Universidad Aut\'onoma de Madrid,
Cantoblanco E-280049, Spain}

\date{\today}

\begin{abstract}
Recent observations constrained the tangential velocity of M31 with
respect to the Milky Way (MW) to be $v_{\rm M31,tan}<34.4$ \kms and
the radial velocity to be in the range $v_{\rm M31,rad}=-109\pm
4.4$\kms \citep{vanderMarel12}. In this study we use a large volume
high resolution N-body cosmologicalsimulation (BOLSHOI) to
statistically study this kinematics in the context of the $\Lambda$CDM
cosmology and three constrained simulations designed to reproduce the
Local Group (LG) and the local cosmic neighborhood. A LG is defined
here as  composed by two dominant separate halos hosting the MW and
the M31 galaxy fulfilling certain isolation criteria. An ensemble  of
all BOLSHOI pairs fulfilling this selection is constructed, and also a
subsample of it which obeys  more stringent observational constraints
on the mass and separation of of the MW and M31 (hereafter the full-
and sub- samples). The comparison of the ensembles of simulated pairs
with the observed LG has been done with respect to the observed radial
and tangential,  the reduced orbital energy ($e$) and angular momentum
($l$) of the LG, and with respect to the dimensionless spin parameter,
$\lambda$. Our main results are: 
%
%
% Jaime here I suggest to give in a very brief way the numbers of Table 2 that I have suggested.
%
The three constrained LGs are more consistent with the unconstrained
sample of LGs rather then with the observed LG.

It follows that the quasi-conserved quantities quantities that
characterize the orbit of the LG, i.e. $e$, $l$ and $\lambda$, do not
challenge the standard $\Lambda$CDM model, but that model is in
tension with regard to the particular phase on the orbit of the LG,
namely the actual values of the radial and tangential velocities. This
might hint to a problem of the $\Lambda$ CDM model to reproduce the
observed LG.

\end{abstract}

\begin{keywords}
{galaxies: kinematics and dynamics, Local Group, methods:numerical}
\end{keywords}

\section{Introduction}

The Milky Way (MW) and Andromeda galaxy (M31) are the dominant
galaxies in the Local Group (LG). Astronomical observations of their
mass distribution impose constraints on the standard cosmological
model. The satellite overabundance problem \citep{Klypin99,Moore99},
tidal disruption features \citep{pandas09} and the disk dominated
morphology \citep{Kazantzidis2008} are examples on of how LG studies
are linked to the cosmological context. Detailed studies on the
Magellanic Clouds dynamics and their possible link to M31 add to the
interest of understanding the details of the LG kinematics and
dynamics
\citep{Besla2007,Tollerud2011,Knebe2011,Fouquet2012,Teyssier2012}. However,
a general concern in the use of the LG as a tool for near-field
cosmology \citep{Freeman2002,Peebles2010} is how typical is the LG
regarding the properties of interest
\citep{Busha2011,Liu2011,ForeroRomero2011,Purcell2012}.

A new valuable piece of information in this issue is the recent
observational determination of the proper-motion measurements of M31,
which until recently had been out of reach \citep{vanderMarel12}.  The
reported measurements set an upper bound for the tangential velocity
of M31 with respect to the MW of $v_{\rm tan,M31}\leq 34.4$ km
s$^{-1}$. Together with the values of the relative radial velocity of
$v_{\rm rad,M31}=-109\pm 4.4$ \kms observations show that the relative
motion of the MW and Andromeda is consistent with a head-on
collision. With this information it is possible to quantify how common
is this kinematic configuration in a $\Lambda$CDM Universe.

This Ltter presents such study. We use a large volume, high resolution
dark matter only N-body simulation in the concordance $\Lambda$CDM
cosmology to find a set of halo pairs with similar  characteristics as
inferred in the LG. We quantify these results in terms of the number
of pairs with given radial and tangential velocities in the
galactocentric rest frame. We also find the pairs that are consistent
with a head on collision in terms of the ratio of the radial to
tangential velocity $f_{\rm t}\equiv v_{\rm tan}/v_{\rm rad}<0.3$ and
present these results in terms of the reduced angular momentum and
mechanical energy.

In addition we make use of three  constrained N-body simulations which
are constructed to reproduce the observed large scale structure of the
Local Universe on scales of a few tens of Mpc. The special feature of
these simulations is that each volume features a pair of halos with
the right characteristics to be considered LG-like objects.

This Letter is structured as follows. In the next section we present
the N-body simulations, the criteria we use to select LG-like halo
pairs. In Section \ref{sec:results} we present the results for the
dynamics in these pairs in terms of the tangential/radial velocities
and the orbital angular momentum/mechanical energy. In the same
section we sumarize these dynamical results in terms of the
dimensionless spin parameter of the pairs. Finally, in the last
section we comment and conclude about the implications of these
results in the context of the $\Lambda$CDM model.

\label{subsec:lg-sample}
\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/separation_BDM.pdf}
\includegraphics[keepaspectratio=true,width=0.45\textwidth]{./figures/total_mass_BDM.pdf}
\caption{\label{fig:distros} Distributions of separations and total
  halo mass (Milky Way and M31) in the LG-like pairs selected from the
  Bolshoi Simulation.} 
\end{center}
\end{figure*}

\section{Simulation and Environment}
\label{sec:methods}
\subsection{The Bolshoi and Constrained Simulations}

The Bolshoi simulation follows the non-linear evolution of the dark
matter density field using N-body techniques. The simulation has a
cubic volume of $250$\hMpc comoving on a side, sampled with $2048^{3}$
particles. The cosmological parameters used in the simulation are
$\Omega_{m}=0.27, \Omega_{\Lambda}=0.73, \sigma_{8}=0.82, h=0.70$ and
$n=0.95$, corresponding to the matter density, vacuum energy density,
the normalization of the power spectrum, the dimensionless Hubble
constant and the index of the slope in the initial power
spectrum. This set of parameters is compatible with the analysis  of
the seventh year of data from the Wilkinson Microwave Anisotropy Probe
(WMAP) \citep{Jarosik2011}. A detailed description of this simulation
can be found in \citep{Bolshoi}.

With these parameters the mass per particle is $m_{p}=1.4\times
10^{8}$\hMsun. In this paper we use algorithms obtained through the
Bound Density Maxima (BDM) algorithm \citep{KlypinBDM}. The halos are
selected to have an overdensity of 200 times the critical
density. Furthermore, we only include in the analysis haloes whose
center is located outside the virial radius of any other halo.  We
have obtained the data through the public available Multidark
database \footnote{{\tt http://www.multidark.org/MultiDark/}}
\citep{2011arXiv1109.0003R}. The database allows us to obtain the
comoving positions, peculiar velocities and masses for all the halos
in the simulation volume at $z=0$. The positions and velocities of
these haloes correspond to the average values of the $250$ most bound
particles. The Hubble flow is taken into account to convert the
peculiar velocities into physical velocities and allow for a
comparison with observations. We have verified that the main
conclusions of this paper hold in the case of halos defined by a FOF
algorithm with a linking lentgth 0.17 times the interparticle
distance.

The constrained simulations we use in this Letter are part of the
Constrained Local UniversE Simulations (CLUES) project whose main
objective is to reproduce the large scale structure in the Local
Universe as accurately as possible. The algorithm and observational
constraints to construct the initial conditions are  described in
\citep{clues2010}.  We use three dark matter only simulations, each
has a cubic volume of $64$\hMpc on a side, with the density field
sampled with $1024^3$ particles. The cosmological density parameter is
$\Omega_m=0.28$, the cosmological constant $\Omega_{\Lambda}=0.72$,
the dimensionless Hubble parameter $h=0.73$, the spectral index of the
primordial density perturbations $n=0.96$ and the power spectrum
normalization $\sigma_{8}=0.817$, also consistent with WMAP 7th year
data.

\subsection{The LG-sample}
Based on the BDM catalogs in the Bolshoi simulation we construct a
halo pair sample with the dynamical  properties consistent with  those
of  the MW and M31. The  criteria  we impose to define a LG-like halo
pair are the following:

\begin{enumerate}
\item Each halo has a mass in the range $5\times10^{11}\hMsun
  <M_{h}<5\times 10^{12}\hMsun$. 
\item With respect to each halo, there cannot be any other halo within
  the mass range $5\times10^{11}\hMsun <M_{h}<5\times 10^{12}\hMsun$
  closer than its partner. It means that there cannot be ambiguity on
  the identity of the pair members. 
\item The relative radial velocity between the two halos is negative
  \citep{vanderMarel12}. 
\item The distance between the center of mass of the halos must be
  less than $0.7$\hMpc \citep{ribas05,vanderMarel08}. 
\item There cannot be halos more massive than $5\times 10^{12}$\hMsun
  within a radius of $2$\hMpc with respect to every object centre
  \citep{Karachentsev04,Anton09}. 
\item There cannot be halos more massive than $5\times 10^{13}$\hMsun
  within  a radius of $3$\hMpc with respect to every object centre
  \citep{Karachentsev04}. 
\end{enumerate}

%YH: Jaime, aren't you using a Virgo constraint?
%J.: What is a Virgo constraint in terms of Bolshoi? I am using the same constraints that we used in the old DMassembly paper.

This sample in the Bolshoi simulation has $1923$ pairs. Additionally,
there is a sample of three (3) pairs constructed from the three
constrained realizations. These pairs fulfill all the above mentioned
conditions and additionally are located in a place with the right
distances with respect to the Virgo cluster in the simulation. Figure
\ref{fig:distros} show the  distribution of separations and total pair
mass computed from the $1923$ pairs in the Bolshoi simulation.

The full observational characteristics that we take in this Letter for
the MW-M31 pair are listed in Table \ref{table:1}.  
A more restrictive subsample of LG-like objects has been constructed
so as to fit obey the observational bounds on the masses and
separation of the two main halos. These amount to:   
\begin{enumerate}
\item The separation between the center of mass of the halos is in the range $700-800$kpc \citep{ribas05,vanderMarel08}.
\item The total mass of the two halos is in the range $1-4\times 10^{12}$\Msun \citep{vanderMarel12}.
\end{enumerate}

Including these conditions the full pair sample is reduced from $1923$
to $158$ pairs. Note that only one constrained LG-like object is
included in the subsample.

\begin{table}
\caption{Summary of observational constraints. The kinematic
  properties for M31 are reported in the galactocentric
  \citep{vanderMarel12}. Values in parenthesis correspond to vector
  components. $\sigma_{\bf x}$ represents the uncertainty on the
  components of vector ${\bf x}$. The values for the individual halo
masses are consistent with the priors used by \cite{vanderMarel12}.}
\begin{center}
\begin{tabular}{ccc}\hline
$v_{\rm M31,rad}$ &(\kms) & $-109.3\pm 4.4$\\
$v_{\rm M31,tan}$ &(\kms) & $<34.4$\\
$r_{\rm M31}$ &(kpc) & $770\pm 40$\\
${\bf r}_{\rm M31}$ & (kpc) &$(-378.9, 612.7, -283.1)$\\
$\sigma_{{\bf r}, {\rm M31}}$ & (kpc) &$(-18.9, 30.6, 14.5)$\\
${\bf v}_{\rm M31}$ & (\kms) & $(66.1, -76.3, 45.1)$\\
$\sigma_{{\bf v},{\rm M31}}$ & (\kms) &$(26.7, 19.0, 26.5)$\\
$M_{\rm 200, MW}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $1.6\pm0.5$ \\
$M_{\rm 200,MW} + M_{\rm 200, M31}$ & ($10^{12}\Msun$) & $3.14\pm 0.58$\\
$\log_{10}\lambda$& & $-1.72\pm 0.07$ \\\hline
\end{tabular}\\
\end{center}
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}.
Note: the value for $\log_{10}\lambda$ is obtained in this Letter
from a Monte Carlo simulation as described in Section XX.
\label{table:1}
\end{table}

\begin{table}
\caption{Summary of results from $\Lambda$CDM.}
\begin{center}
\begin{tabular}{cccc}\hline
 &  & Full Sample & Reduced Sample\\
$v_{\rm M31,rad}$ &(\kms) & $-70\pm 10$ & $-90\pm 10$\\
$v_{\rm M31,tan}$ &(\kms) & $50\pm 10$ & $50\pm 10$\\
$\log_{10}\lambda$& & $-1.47\pm 0.13$& \\ 
\end{tabular}
\end{center}
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}

\label{table:2}
\end{table}

\section{Results}
\label{sec:results}

\begin{table*}
\caption{Summary of the comparison against $\Lambda$CDM.}
\begin{center}
\begin{tabular}{ccccc}\hline
Physical & (\%) Pairs consistent & \% Pairs consistent & \% Pairs
consistent& (\%) Pairs consistent\\ 
& with observations & with observations & with LCDM & with LCDM\\ 
 & in full sample & in reduced sample & in full sample & in reduced sample\\ \hline
$v_{\rm r}$-$v_{\rm t}$ & (0.4\%) 8/1923 & (0\%)0/158 & (1\%)23/1923 & (8\%)13/158\\
$e_{\rm tot}$-$l_{\rm orb}$ & (15\%)298/1923 & (24\%)38/158 & - & -\\
$\log_{10}\lambda$ & (13\%)257/1923 & (9\%)15/158 & - & -\\
$r_{\rm t}=v_{\rm t}/v_{\rm r}$& (12\%)242/1923 & (8\%)13/158 & -& -\\
\end{tabular}
\end{center}
\vspace{1mm}
Note: the observational uncertainties in the position vector
correspond to a $5\%$ in each component consistent with the
uncertainties in the distance \citep[see references
  in][]{vanderMarel08}

\label{table:2}
\end{table*}

\subsection{Radial and tangential velocities}

Figure \ref{fig:rt} summarizes the central finding of this
Letter. Most of the LG-like pairs (left panel) in the Bolshoi
Simulation and the pairs in the constrained simulation have radial and
tangential velocities notably different from the observational
constraints.  The most probable velocities in the simulation are
around $v_{\rm rad,\Lambda CDM} = -70\pm5$\kms and $v_{\rm tan,\Lambda
  CDM} = 50\pm5$ \kms, where the uncertainties in these values reflect
the minimum grid size needed to obtain robust statistics from the 2D
histogram. There are $23$ pairs out of the full LG-like pairs sample
centered in those values within the same range of absolute
uncertainties allowed by observations (i.e. $\sigma_{\rm tan}=17$\kms
and $\sigma_{\rm rad}=4$\kms.) In contrast, there are only $8$ pairs
in the interval allowed by observations.

The left panel in the same Figure \ref{fig:rt} shows the results for a
more restrictive selection of objects in the LG-like sample, reducing
the number of objects from  $1923$   to $158$ and their most probable
configuration is $v_{\rm rad,\Lambda CDM} = -50\pm5$\kms and $v_{\rm
  tan,\Lambda CDM} = 50\pm5$\kms with $5$ objects within that interval
and none consistent with observations. In this case none of the pairs
are found in the interval allowed by observations. The high eccentric
orbit of the observed LG  constitutes an unlikely configurations for
the $\Lambda$CDM LG-like objects. This holds for the full and the
subsample of objects.

From Figure \ref{fig:rt} it is also clear that there is a significant
number of pairs with a high tangential-to-radial velocity ratio. The
peak in the pair number density is located around a region of $f_{\rm
  t}\equiv v_{\rm tan \Lambda CDM}/v_{\rm rad \Lambda CDM}\sim 0.7$,
while the observations suggest $f_{\rm t}<0.32$. In the Bolshoi
Simulation, out of $1923$ LG-like pairs we find that $242$ pairs
($\sim 12\%$ of the sample) have values for $f_{\rm t}$ consistent
with that constraint, for the more restrictive sample there are $13$
pairs within that constraint ($\sim 8\%$ of the sample). The three
pairs from the constrained realizations are also outside the region of
the most probable values in a $\Lambda$CDM cosmology with
tangential-to-radial ratios of $f_{\rm t}= 0.35, 0.45, 0.73$.

The difference between the left and right panels of Figure
\ref{fig:rt} reflects  the tension found from the full and the sub-
sample of the LG-like objects. {\bf Yehuda, What tension??? The
  tension comes between theory and observation, not between the two
  theoretical results.}

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM.pdf}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_rt_BDM_narrow.pdf}
\caption{Histograms of the radial and tangential velocities for
  LG-like halo pairs in the Bolshoi simulation. The left panel
  corresponds to the general LG-like pair simple, the right panel
  corresponds to the subsample with cuts on pair separation and total
  mass, matching the observational constraints. The squared panel
  shows as a shaded histogram the number of pairs on the radial and
  tangential velocities plane. The rectangular panels show the
  histograms when only one of the velocity components is used to bin
  the data. The half ellipse corresponds to the observed observational
  constraints for M31 in the Galactocentric rest frame: $v_{\rm
    rad}=109.2\pm 4.4$ km s$^{11}$ and $v_{\rm tan}< 34.4$ km s$^{-1}$
  \citep{vanderMarel12}. The circles represent the positions of the
  pairs from the constrained simulations. The highest number density
  in the left panel is obtained around $v_{\rm rad,\Lambda CDM} =
  -70\pm5$ \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ \kms, for the
  right panel it is located around $v_{\rm rad,\Lambda CDM} = -50\pm5$
  \kms, $v_{\rm tan,\Lambda CDM} = 50\pm5$ } 
\label{fig:rt}
\end{center}

\end{figure*}

\subsection{Reduced Angular Momentum and Energy}

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM.pdf}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_EJ_BDM_narrow.pdf}
\caption{Histograms of the orbital angular momentum ($l_{\rm orb}$)
  and mechanical energy ($e_{\rm tot}$) per unit of reduced mass
  calculated considering the halos as point masses. The panel
  distribution is the same as in Figure \ref{fig:rt}. The contour line
  encloses $68\%$ of the Monte Carlo generated points to estimated the
  uncertainties from the observational values summarized in Table
  1. The constraints in this plot are less restrictive than in
  Fig.\ref{fig:rt} due to added uncertainties on the tangential
  velocity ($100\%$), the square of the norm of the velocity ($40\%$)
  and the total mass of the two halos ($20\%$). } 
\label{fig:EJ}
\end{center}
\end{figure*}

The two-body problem of point-like masses can serve as a proxy for the
dynamics of the LG and be used as a tool for  studying the LG within
the framework of the standard cosmological model. Within the model the
dynamics of the LG is governed by the (center of mass)  reduced
angular momentum $l= {{\bf r}_{\rm M31}}\times{\bf v}_{\rm M31}|$ and
reduced energy  $e=\frac{1}{2} {\bf v}_{\rm M31}|^2 - G M/|{\bf
  r}_{\rm M31}|$,  where the total mass is $M=m_{\rm M31}+m_{\rm MW}$,
the reduced mass is $\mu=m_{\rm M31}m_{\rm MW}/M$ and $G$ is the
gravitational constant.

This formulation has a clear theoretical advantage if one considers
the angular momentum and the mechanical energy as quasi-conserved
dynamical quantities. This means that, after some formation time,
these quantities do not significantly evolve as it is the case for the
radial and tangential velocities.

Conversely, there is an observational disadvantage. These quantities
are derived from very different kinds of observations, increasing the
uncertainties in their final determination.

A possible conflict between the likelihood of the LG in the
$\Lambda$CDM model as viewed from the radial and tangential velocities
as opposed to the angular momentum and energy perspectives would point
to a particular phase on the LG two-body orbit. Given the masses,
velocity and position of the observed MW and M31 (TAble \ref{table:1}
) the reduced energy and angular momentum are calculated. The Monte
Carlo sampling is used to to estimate the uncertainty in these
quantities.

Figure \ref{fig:EJ} presents the reduced  energy and angular momentum
distribution of the constrained and unconstrained LG-like  objects of
the full and subsample, following the   plotting conventions from
Figure \ref{fig:rt}.  The  1-$\sigma$ contour of Figure \ref{fig:EJ}
does not cross the zero level, as can be naively expected from Figure
\ref{fig:rt}. This does happen for the $\approx 80\%$ level
contour. The small difference between the two presentations of the
observational uncertainties stems from the Monte Carol sampling.

Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and
$l_{\rm orb}$ constraints are less restrictive compared with the
radial and tangential velocity constraints. For the unconstrained full
(sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$
observational constraints. All the constrained LGs of the full
samples are consistent with the 1-$sigma$ observational uncertainty.

However a reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
bring these pairs and the $\Lambda$CDM expectation outside the
1-$\sigma$ observational uncertainty.

\begin{figure*}
\begin{center}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM.pdf}
\includegraphics[keepaspectratio=true,width=0.46\textwidth]{./figures/test_lambda_BDM_narrow.pdf}
\caption{{\rm \label{fig:lambda} Normalized histograms of Peeble's
    spin parameter $\lambda$ for the pairs in the Bolshoi simulation
    and its infered values for the LG from the observational
    constraints from a Monte Carlo simulation. The left panel includes
    all the LG-pairs while the right panel includes a reduced sample
    of pairs that match the separation and total mass infered from
    Local Group observations. The vertical lines with the white dots
    represent the values inferred for pairs in the constrained
    simulations.}} 
\label{fig:lambda}
\end{center}
\end{figure*}

\subsection{Dimensionless Spin Parameter}

The dimensionless 
spin parameter $\lambda$ \citep{Peebles1971} is used here to
characterize the dynamical state of the observed and simulated
LGs. The parameter measures the dynamical role of the angular momentum
in terms of the gravitational attraction and is defined by

\begin{equation}
\lambda = \frac{\mu^{3/2} l_{\rm orb} \sqrt{e_{\rm tot}}}{G M^{5/2}}, 
\end{equation}

where the $\mu=m_{\rm M31}m_{\rm MW}/M$ is the reduced mass for the pair.

We compare the distribution for  $\lambda$ obtained from the pair
population in the Bolshoi simulation and the distribution from the  Monte Carlo
simulation used to estimate the uncertainties on $e_{\rm tot}$ and $l_{\rm orb}$. 
Figure \ref{fig:lambda} shows the likelihood distribution of the
observational errors and the spin parameter of the LG-like pairs of
the full (left panel) and the sub- (right panel) samples.  The value
of $\lambda$ of the constrained LGs is shown as well.

The expected configuration in energy and angular momentum of the LG is
slightly inconsistent with the expectation from halo pairs in $\Lambda
CDM$.

The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value. 
Inspection of Figure \ref{fig:lambda} shows that the $\lambda$
distribution is close to a lognormal one and thus resembles the
distribution of the spin parameter of DM halos in simulation ({\bf
  Jaime - please provide a reference}) .  
The estimated value of $\log_{10}\lambda$ of the observed LG is
$-1.76$, compared with the mean value of $\log_{10}\lambda$ of $-1.51$
for the full sample and $-1.38$ for the subsample. The values of the
spin parameter of the 3 constrained LGs are $\log_{10}(\lambda_{\rm
  CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging
to the sole constrained LG of the subsample.

It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology. 
%JF. Yehuda, I don't understaind this sentence. 
%This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
%YH

\section{Conclusions}
We have presented a comparison between the observed kinematics for the
M31 in the galactocentric restframe and the expectations for a large
N-body cosmological simulation in the $\Lambda$CDM cosmology. In the
simulation we select a sample of halo pairs in the mass range $5\times
10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation
conditions of the Local Group from other massive structures. While the
observations show that M31 moves towards us  on a highly eccentric orbit,
 the simulation shows that the most common configuration at $z=0$ has
 values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan,
   \Lambda CDM}=62\pm5$ \kms.

Using the same absolute values for the uncertainty in the observed
velocity components, we find that halos within the preferred
$\Lambda$CDM values are five times more common than pairs compatible
with the observational constraint.  The qualitative nature of these
results is still valid after a narrower selection on separation and
total pair mass. Additionally, pairs with a fraction of tangential to
radial velocity $f_{\rm t}<0.32$ (similar to observations) represent
$8\%$ of the total sample of LG-like pairs. Making an tighter
selection to match the observational constraints on the separation and
total mass results in zero pairs compatible with observations.

Approximating the LG as two point masses we express the above
mentioned results in terms of the orbital angular momentum $l_{\rm
  orb}$ and the mechanical energy $e_{\rm tot}$ per unit of reduced
mass. We find that the uncertainties in the tangential velocity, the
square of the norm of the velocity and the total mass in the LG
produce poorer constraints on the number of simulated pairs that are
consistent with the observations. Nevertheless, in the case of the
LG-pair sample that also fulfills the separation and total mass
criteria there is a slight tension between simulation and
observation. A reduction by a factor of $2$ in the observational
uncertainty on the radial velocity would clarify this issue.

In the three pairs from constrained simulations we find kinematics
dominated by radial velocities. However their velocity components
differ from the observational constraints and their mechanical energy
and orbital angular momentum are in broad concordance with
observations. There is only one pair that fullfills all the
separation, total mass constraints and matches the most probable value
for the dimensionless spin parameter $\lambda$ inferred from
observations.

The $\lambda$ spin parameter is itself a useful instrument to gauge
the dynamical state of pairs. Reduced uncertainties in masses,
distances and velocities can be included to produce an updated
estimate for this parameter. This will immedialtly improve the
statistical comparison between our Local Group and the expectations
from $\Lambda$CDM.

In this Letter we have shown that LG-like pairs in $\Lambda$CDM show
prefered values for their relative velocities, angular momentum and
total mechanical energy. The values for the orbital angular momentum
and energy, merged into the $\lambda$ spin parameter, are in mild
disagreement with the observational constraints. However, there is a
strong tension with the precise values for the radial and tangential
velocities.

Under the approximation of consevartion of the orbital angular
momentum and mechanical energy, we see that an agreement in the total
angular momentum and energy and a marked difference with the precise
balance between today's radial and tangential velocities could only be
explained if the initial conditions for the formation of the Local
Group are special in comparison to the initial conditions of any other
pair of dark matter halos in the $\Lambda$CDM cosmology.

We consider that reducing the uncertainties in the observations
relevant to this study are needed. Likewise, additional theoretical
work to understand the origin and implications of this result deserve
further study.

\label{sec:conclusions}
\section*{Acknowledgments}  
J.E.F-R acknowledges financial support from the Vicerrector\'{\i}a de
Investigaciones at Universidad de los Andes through its {\it Fondo de
  Apoyo a Profesores Asistentes} and the Peter and Patricia Gruber
Foundation through its fellowship administered by the International
Astronomical Union. J.E.F-R also acknowledges early discussions with
Alejandro Garc\'ia that motivated this work.

The data and source code and instructions to replicate the results of
this paper can be found as a github repository {\texttt
  https://github.com/forero/LG\_Kinematics/}. Thanks to the IPython
community \citep{IPython}. Thanks to Jessica Kirkpatrick for releasing
her Python code to make nice plots of 2D histograms.

The MultiDark Database used in this paper  was constructed as part of
the activities of the German Astrophysical Virtual Observatory as
result of a collaboration between the Leibniz-Institute for
Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider
Project CSD2009-00064. The Bolshoi simulation was run on the NASA's
Pleiades supercomputer at the NASA Ames Research Center.

\bibliographystyle{apj}
\bibliography{references}

\end{document}