### 2013-03-01 paper/reply_referee.txt
referee reply

Dear Editor,

We have appreciated the positive referee's comments and the suggestions
for improvement in the paper.

In what follows below the dashes, the referees comments are indented
and our responses are not commented.

Jaime E. Forero-Romero (on behalf of the authors)

----

	First, it would be useful if the authors were more explicit about how
	they are defining "consistency". Reading the text fairly carefully, I
	think that I was able to discern that you are using 1-sigma
	constraints on v_tan to define consistency, but this needs to be
	stated very clearly. For example, what is the quoted c.l. for the HST
	limit on V_tan listed in Table 1? This should be stated clearly in the
	table itself. In the caption of Table 3, again describe exactly how
	you are defining consistency. It would be useful to know how these
	numbers change as you very the definition (1-sigma, 2, 3?). 
	I discern from a footnote that the contours drawn in Figure 1 are
	1-sigma contours. Please state this clearly in the caption
	and/or in the text that describes the figure.

We have now included explicit mention of the 1-sigma criteria in Table
1, Table 3 and the caption of Figure 1.

   	 Second, the authors make a good case pairs that have Vr/Vtan values
	 similar to those observed are somewhat rare in the simulations and
	 suggest that this might possibly imply that the LG is not typical. It
	 should be possible to ask some basic questions about the small # of
	 systems that appear consistent with the data: does it matter? Do they
	 have outlying halo spins, concentrations, or formation times? What
	 about their larger-scale environments?

The answers to these questions are being clarified in a paper under
preparation by one of the coauthors (S. Bustamante).

The largest difficulty to study the pairs consistent with Vr/Vtan is
their low number. There are only 8 of them for the general sample and
_none_ in the reduced sample. This makes any statistical claim hard to
obtain. However we have found that these pairs prefer special large
scale environments defined by the tidal and velocity shear tensor
fields, namely they live in filaments. We are also investigating their
alignment with respect to the directions along this filaments. We will
include in that study the referee sugestions to study the
concentration and spin parameter.

We consider that this low number matters to the extent that it shows
how unlikely is to have Local Group like kinematic configurations in
LCDM, which might be relevant to explain other features in the matter
density field such as the special alignment of satellites in the Milky
Way and Andromeda or might hint towards a constraint on the expected
MW assembly history in LCDM.